---
title: "R Project 3"
author: "Huy Huynh"
date: "2/26/2020"
output:
  html_document: default
  pdf_document: default
---


Note that these exercises should be performed using `dplyr` (do not directly access or manipulate the data frames). 

Turn in the html to blackboard with the filename `RProject3_yourlastname.html`. **Points will be taken off for not doing this.** 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Part 1 

Install and load the "fueleconomy" package and the tidyverse

```{r}
# run these only once
# install.packages("devtools")
# devtools::install_github("hadley/fueleconomy")
```

```{r}
library(tidyverse)
library(fueleconomy)
glimpse(vehicles)
```



1. Select the different manufacturers (makes) of the cars in this data set. Save this vector in a variable
```{r}
makes <- select(vehicles,make)
```


2. Use the `distinct()` function to determine how many different car manufacturers are represented by the data set

```{r}
nrow(distinct(makes))
```

3. Filter the data set for vehicles manufactured in 1997

```{r}
vehicle_1997 <- filter(vehicles, year == 1997)
vehicle_1997
```

4. Arrange the 1997 cars by highway (`hwy`) gas milage
```{r}
vehicle_1997 <- arrange(vehicle_1997, hwy)
vehicle_1997
```


5. Mutate the 1997 cars data frame to add a column `average` that has the average gas milage (between city and highway mpg) for each car

```{r}
vehicle_1997 <- mutate(vehicle_1997, average = (hwy + cty) / 2)
vehicle_1997
```

6. Filter the whole vehicles data set for 2-Wheel Drive vehicles that get more than 20 miles/gallon in the city.  Save this new data frame in a variable.
```{r}
two_wheel <- filter(vehicles, drive == "2-Wheel Drive", cty > 20)
```


7. Of the above vehicles, what is the vehicle ID of the vehicle with the worst hwy mpg? Hint: filter for the worst vehicle, then select its ID.
```{r}
filtered <- filter(two_wheel, hwy == min(hwy))
worst_hwy <- select(filtered, id)
```


8. Write a function that takes a `year_choice` and a `make_choice` as parameters, and returns the vehicle model that gets the most hwy miles/gallon of vehicles  of that make in that year. You'll need to filter more (and do some selecting)!

```{r}
make_year_filter <- function(make_choice, year_choice) {
  filtered <- filter(vehicles, make == make_choice, year == year_choice)
  filtered <- filter(filtered, hwy == max(hwy))
  selected <- select(filtered, model)
  selected
}
```

9. What was the most efficient Honda model of 1995?
```{r}
make_year_filter("Honda", 1995)
```

10. Which 2015 Acura model has the best hwy MGH? (Use dplyr, but without method
chaining or pipes--use temporary variables!)
```{r}
acuras <- filter(vehicles, make == "Acura", year == 2015)
best_acura <- filter(acuras, hwy == max(hwy))
best_model <- select(best_acura, model)
best_model
```


11. Which 2015 Acura model has the best hwy MPG? (Use dplyr, nesting functions)

```{r}
best_model <- select(
  filter(
    filter(vehicles, make == "Acura", year == 2015), hwy == max(hwy)
  ), model
)
best_model
```

12. Which 2015 Acura model has the best hwy MPG? (Use dplyr and the pipe operator)

```{r}
best_model <- filter(vehicles, make == "Acura", year == 2015) %>%
  filter(hwy == max(hwy)) %>%
  select(model)
best_model
```

## Bonus

Write 3 functions, one for each approach in 10, 11 and 12.  Then, test how long it takes to perform each one 1000 times
```{r}

```


# Part 2 


Read in the data (from `pulitzer-circulation-data.csv`). Remember to 
not treat strings as factors!

```{r}
pulitzer <- read.csv("pulitzer-circulation-data.csv", stringsAsFactors = FALSE)
```

1. View in the data set. Start to understand what the data set contains

```{r}
view(pulitzer)
```

2. Print out the names of the columns for reference

```{r}
colnames(pulitzer)
```

3. Use the 'str()' function to also see what types of values are contained in each column (you're looking at the second column after the `:`) Did any value type surprise you? Why do you think they are that type?

```{r}
str(pulitzer)
```
Daily.Circulation of 2004 and 2013 and Change.in.Daily.Circulation..2004.2013 did surprise me since I beleive these should be integers instead of characters. 

4. Add a column to the data frame called 'Pulitzer.Prize.Change` that contains the difference in the number of times each paper was a winner or finalist (hereafter we'll call this group "winner") during 2004-2014 and during 1990-2003
```{r}
mutate(pulitzer,
  Pulitzer.Prize.Change =
    Pulitzer.Prize.Winners.and.Finalists..2004.2014- Pulitzer.Prize.Winners.and.Finalists..1990.2003
)
```


5. What was the name of the publication that has the most winners between 2004-2014?
```{r}
filter(pulitzer, max(Pulitzer.Prize.Winners.and.Finalists..2004.2014) ==
  Pulitzer.Prize.Winners.and.Finalists..2004.2014) %>%
  select(Newspaper)

```


6. Which publication with at least 5 winners between 2004-2014 had the biggest decrease(negative) in daily circulation numbers?
```{r}
filter(pulitzer, Pulitzer.Prize.Winners.and.Finalists..2004.2014 >= 5) %>%
  filter(min(Change.in.Daily.Circulation..2004.2013) == Change.in.Daily.Circulation..2004.2013) %>%
  select(Newspaper)
```


7. An important part about being a data scientist is asking questions.  Write a question you may be interested in about this data set, and then use   dplyr to figure out the answer!

```{r}
#What was the name of the publication that has the least winners between 1990-2014?
filter(pulitzer, min(Pulitzer.Prize.Winners.and.Finalists..1990.2014) ==
  Pulitzer.Prize.Winners.and.Finalists..1990.2014) %>%
  select(Newspaper, Pulitzer.Prize.Winners.and.Finalists..1990.2014 )
```

 